---
title: "co kriging Italica"
author: "Esperanza"
date: "6/5/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=TRUE, warning=TRUE}
library(rgdal)
library(raster)
library(ggspatial) ##ggplot for spatial analysis
library(ggplot2)
library(GISTools)
library(maptools)
library(spdep) ##autocorrelation analysis
library(pgirmess) 
library(gstat)
library(scales)
library(car)
library(MASS)
```


Import dataset and access the data.

```{r message=FALSE, warning=FALSE}
setwd("C:/TESIS/Output/co_kriging_v2")
cb0= readOGR ("C:/TESIS/Output/co_kriging_v2/c0_base.shp")
grd1= readOGR ("C:/TESIS/Output/co_kriging_v2/grid.shp")

```

```{r}
cb0.df=as.data.frame(cb0)
```



# Step 1: co-kriging

Inspect the distribution of values through histogram.

```{r}
hist(cb0$Z)
```


```{r}
cb0.df= data.frame(cb0)
```

The assessment confirms that we don't need to normalise the data.

## Moran's I Correlogram - Area 1

```{r}
cb0.cor= correlog(coordinates(cb0), cb0$Z, method = "Moran")
cb0.cor

```
```{r}
cb0.cor= as.data.frame(cb0.cor)
ggplot(cb0.cor, aes(dist.class,coef))+
  geom_point(aes(col=p.value), size= 3)+
  labs(x = "Distance classes", y = "Moran's I Statistic",
  colour = "P Value",  title = "Correlogram")
```

## Local Moran's I (Anselin's LISA)

0-100 lag will be used based on the correlogram.

```{r}
cb0.dnn=dnearneigh(cb0,0,100)
summary(cb0.dnn, cb0, longlat=TRUE, scale=0.5)
```

1 region within this distance lag have no neighbours. So we try with the second distance lag: 0-200

```{r}
cb0.dnn=dnearneigh(cb0,0,200)
summary(cb0.dnn, cb0, longlat=TRUE, scale=0.5)
```

All the regions have neighbours, so this distance lag will be used to compute Local Moran's I. 

```{r}
cb0.dnn.w<-nb2listw(cb0.dnn,style="B") 
cb0.lmoran<-localmoran(cb0$Z,cb0.dnn.w) 
```

Now we plot the result of the the local Moran's I. 

```{r}
pval=ifelse(cb0.lmoran[,5]<0.05,"p< 0.05","p> 0.05")
I=as.vector(cb0.lmoran[,1])
ggplot()+
  layer_spatial(cb0,aes(size=I, col= pval))+
  labs(size = "Moran's I",  col = "P Value")
```

## Variogram

```{r}
max(dist(coordinates(cb0)))
cb0.var= variogram(cb0$Z~1,cb0, cutoff=500)
plot(cb0.var)
cb0.fitvar= vgm(50,"Sph",170,0)
plot(cb0.var,cb0.fitvar)
```

## Ordinary Kriging 

In order to perform Ordinary Kriging we create a grid from grd1.

```{r}
grd1.df= as.data.frame(coordinates(grd1))

coordinates(grd1.df) <- ~coords.x1 + coords.x2
gridded(grd1.df) <- TRUE
```

```{r}
cb0.df= as.data.frame(cb0)
cb0.spdf= SpatialPointsDataFrame(cb0.df[2:3], cb0.df[1])
```

Now we use the krige function and plot the results. 

```{r}
cb0.ok<-krige(cb0$Z~1,cb0.spdf,grd1.df,cb0.fitvar)
cb0.ok %>% as.data.frame %>%
  ggplot(aes(x=coords.x1, y=coords.x2)) + geom_tile(aes(fill=var1.pred)) + coord_equal() +
  scale_fill_gradient(low = "yellow", high="blue") +
  scale_x_continuous(labels=comma) + scale_y_continuous(labels=comma) +
  theme_bw()
ok.ras= raster(cb0.ok)
writeRaster(ok.ras,
            filename = "ok_base", "GTiff", overwrite=T)
```
The prediction seems to be to much influenced by the distribution of points. Hence, we try to use co-Kriging to see if is it possible to obtain a more reliable prediction. 

The new variable that will be used for the prediction is the modern elevation of the analysed area. We import data: "c1_base.shp", a Spatial Point Data Frame that contains elevations of both the Roman (Z) and the Contemporary (mde) periods, measured in the same spots. Moreover, we import data "cp_base.shp", which contains measurements of modern elevation in 156 random distributed spots for which we don't know the ancient elevation. 

```{r message=FALSE, warning=FALSE}
cb1= readOGR ("C:/TESIS/Output/co_kriging_v2/c1_base.shp") 
cp= readOGR ("C:/TESIS/Output/co_kriging_v2/cp_base.shp")
```

Plot modern elevation
```{r}
ggplot()+
  layer_spatial(cp, aes(size=mdt, col= mdt))
```

## Correlation between variables

We plot a scatter plot in order to see if a correlation between the two variables exists. In order to do so, we use the dataset containing both ancient and modern elevation. 

```{r}
cb1.df= as.data.frame(cb1)
ggplot(cb1.df, aes(x=Z, y=mdt)) + 
  geom_point(size=4)
```

We compute the correlation coefficient. 

```{r}
cor= cor(cb1.df, method= "pearson")
head(cor)
```

## Variogram 

We plot and fit the variogram of only modern elevation. 

```{r}
cp.var= variogram(cp$mdt~1,cp)
plot(cp.var)
cp.fitvar= vgm(27, "Sph", 300,0)
plot(cp.var,cp.fitvar)
```

We build a gstat structure containing the two sample sets: "ancient elevation" and "modern elevation"
```{r}

gb=gstat(id="ancient elevation",form= cb0$Z~1,data= cb0)
gb=gstat(gb,id="modern elevation",form=cp$mdt~1,data=cp)

```

We create and plot a cross-variogram of the two variables

```{r}
vcross= variogram(gb)
```
```{r}
plot(vcross)
```

No we add a variogram models to the gstat object and fit it using linear model of co-regionalisation.

```{r}
gb= gstat(gb, id= "ancient elevation", model= cb0.fitvar, fill.all = T) 
```

Now the gstat object contains both the data and the models.
As a further step, we fit all the three variograms using the fit.lmc method (fit linear model of co-regionalization) (Rossiter, 2018)

```{r}
(gb= fit.lmc(vcross,gb, fit.method=6, correct.diagonal = 1.01))
```
```{r}
plot(variogram(gb), model=gb$model)
```

We use the predict function
```{r}

grd1.df= as.data.frame(coordinates(grd1))

coordinates(grd1.df) <- ~coords.x1 + coords.x2
gridded(grd1.df) <- TRUE

crs(grd1.df)=crs(cp)

k.c = predict(gb, grd1.df)

```

```{r}
plot(k.c)
```


```{r}
kc.ras= raster(k.c)
writeRaster(kc.ras,
            filename = "ck_base", "GTiff", overwrite=T)
```

